---
title: "Fuel Consumption and Transmission Type"
author: "Masaaki Inaba"
date: "January 16, 2016"
output:
  pdf_document:
    toc: true
---

# 1. Executive Summary

We work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

- Is an automatic or manual transmission better for MPG
- Quantify the MPG difference between automatic and manual transmissions

# 2. Data Processing

```{r}
data(mtcars)
mtcars$am   = factor(mtcars$am)
mtcars$cyl  = factor(mtcars$cyl)
mtcars$vs   = factor(mtcars$vs)
mtcars$gear = factor(mtcars$gear)
mtcars$carb = factor(mtcars$carb)
```

# 3. Regression Analysis

## 3.1. We start regression analysis based on intial model which inclueds all variables and choose a best model by AIC in a Stepwise Algorithm.

```{r results = 'hide'}
initial_model = lm(mpg ~ ., data = mtcars)
best_model = step(initial_model)
```

##3.2. The best model is as below:

```{r}
summary(best_model)
```

# 4. Conclusion

## Question 1. Is an automatic or manual transmission better for MPG?
=> Automatic transmission is better than manual transmission for MPG.

## Question 2. Quantify the MPG difference between automatic and manual transmissions.
=> Manual transmission get 1.80921 more MPG compared to automatic transmission.

# Appendix

## Figure 1. MPG per Transmission

```{r}
boxplot(mpg ~ am, data = mtcars, main = "MPG per Transmission",
        xlab = "Transmission (0 = automatic, 1 = manual)", ylab = "MPG")
```

## Figure 2. Residual Plot and Diagnostics

```{r}
par(mfrow = c(2, 2))
plot(best_model)
```
